---
title: "Model outline"
author: "Erlend B. Nilsen"
date: "20 9 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
```

## Introduction

Distance sampling methods are widely used for estimating abundance (or density) of wildlife species. In this application, I will develop a new version of the distance sampling model suitable for cases where you have access to i) time series data, ii) the observed individuals can be separated into age- and sex classes, and iii) you have access to auxillary data from marked individuals or DNA samples from non-invasive sampling methods. The model will merge ideas from the litterature on Integrated Population Models (hereafter IPM) and distance sampling methods. During the last ten year, there has been a nice developemnt in the distance sampling litterature, notably with the advent of open population variants. 

The model I devlop here will have wide applications, and will form a flexible basis for further ecological inference. Once the model structure has been established and reported (published), the model framework is ready for use to test ecological hypothesis.  

### Workflow
When developing this model, I will follow this workflow: 
- Write down the model, to make explicit all assumptions. 
- Use the model to estimate adundance and demographic parameters from simulated data
- Wrangle and prepare real data from line transect surveys and marked individuals. 





### Line transect data

First we have to access the line transect data, that will constitute the basic data input for the model. The data are openly available through GBIF, as based on the Event Core data structure. To access data, we can use the following code: 


```{r echo=FALSE}




```







